The 2014 SMT Competition

### 2015-08-04 2014/report/SMTCOMP.tex
Additional analyses

The results of the competition per se were not surprising. The solvers that performed best in previous years continued to do so. Boolector wins those divisions that focus on bit-vector problems; CVC4 has the most breadth of application and wins most of the other divisions, with Yices2 also having a strong showing. Solvers that were optimized with particular decision procedures for particular logics can capture individual wins: AProVE for QF\_NIA; veriT for UFLRA. If Z3 had competed, with the current version entered by the organizers, it would have won many divisions, but not all; CVC4 is 
definitely competitive with it. Table \ref{Table:maintrack} does not show the computation time used by the solvers; however, that data is available in the detailed results and shows that of the leading contenders for broad application across logics, Z3 is the most efficient and is the solver to beat in future competitions.

@@

\begin{table}[H]
\caption{Main track results}
\vspace{-\smallskipamount}
\label{Table:maintrack}
\centering
\newcommand{\win}[1]{\textbf{#1}}
\resizebox{!}{299pt}{\begin{tabular}{|l|r|r|l|}
\hline
Logic &	Solvers &	Benchmarks & Order (winner in bold) \\
\hline
ALIA & 		4 & 	29  & 	[Z3]; \win{CVC4}; veriT; CVC3. \\
AUFLIA & 	4 & 	4   & 	\win{CVC4}; [Z3]; CVC3; veriT. \\
AUFLIRA & 	4 & 	10791 & 	[Z3]; \win{CVC4}; CVC3; veriT. \\
AUFNIRA & 	4 & 	564 & 	[CVC4-with-bugfix]; [Z3]; \win{CVC3}; CVC4. \\
LIA & 		4 & 	46 & 	[Z3]; \win{CVC4}; CVC3; veriT. \\
LRA & 		4 & 	171 & 	\win{CVC4}; [Z3]; CVC3; veriT. \\
NIA & 		3 & 	9 &  	[Z3]; \win{CVC4}; CVC3. \\
NRA & 		3 & 	3747 & 	[Z3]; \win{CVC4}; CVC3. \\
QF\_ABV & 	9 & 	6457 & \win{Boolector-j}; Boolector-d; [MathSAT]; \\
& & & SONOLAR; CVC4; [Z3]; Yices2; \\
& & & Kleaver-STP; Kleaver-portfolio. \\
QF\_ALIA & 	5 & 	97 & \win{Yices2}; SMTInterpol; [Z3]; [MathSAT]; CVC4. \\
QF\_AUFBV & 	4 & 	37 & \win{CVC4}; Yices2; [Z3]; [MathSAT]. \\
QF\_AUFLIA & 	6 & 	610 & \win{Yices2}; [MathSAT]; [Z3] \\
& & & SMTInterpol; CVC4; veriT. \\
QF\_AX & 	5 & 	335 & \win{Yices2}; [MathSAT]; [Z3]; CVC4; \\
& & & SMTInterpol. \\
QF\_BV & 	11 & 	2488 & \win{Boolector}; STP-CryptoMiniSat4; \\
& & & [CVC4-with-bugfix]; [MathSAT]; [Z3]; \\
& & & CVC4; 4Simp; SONOLAR; Yices2; \\
& & & abziz\_min\_features; abziz\_all\_features. \\
QF\_IDL & 	4 & 	1315 & 	[Z3]; \win{Yices2}; CVC4; veriT. \\
QF\_LIA & 	7 & 	4381 & 	[CVC4-with-bugfix]; [MathSAT]; \\
& & & \win{SMTInterpol}; Yices2; [Z3]; veriT; CVC4. \\
QF\_LRA & 	6 & 	1343  & \win{CVC4}; Yices2; [MathSAT]; SMTInterpol; \\
& & & veriT; [Z3]. \\
QF\_NIA & 	4 & 	8327 & 	[Z3]; \win{AProVE}; CVC3; CVC4. \\
QF\_NRA & 	4 & 	10121 & [Z3]; \win{CVC3}; CVC4; raSAT. \\
QF\_RDL & 	4 & 	132 & 	\win{Yices2}; [Z3]; veriT; CVC4. \\
QF\_UF & 	7 & 	4124 & 	\win{Yices2}; veriT; CVC4; OpenSMT2; [Z3]; \\
& & & [MathSAT]; SMTInterpol. \\
QF\_UFBV & 	4 & 	31 & 	\win{Yices2}; [Z3]; [MathSAT]; CVC4. \\
QF\_UFIDL & 	4 & 	311 & 	[Z3]; \win{Yices2}; CVC4; veriT. \\
QF\_UFLIA & 	6 & 	484 & 	[Z3]; \win{Yices2}; CVC4; SMTInterpol; \\
& & & [MathSAT]; veriT. \\
QF\_UFLRA & 	6 & 	1176 & 	[Z3]; \win{Yices2}; [MathSAT]; CVC4; \\
& & & SMTInterpol; veriT. \\
QF\_UFNIA & 	3 & 	7 & 	\win{CVC4}; [Z3]; CVC3. \\
QF\_UFNRA & 	3 & 	32 & 	[Z3]; \win{CVC3}; CVC4. \\
UF & 		4 & 	2830 & 	\win{CVC4}; [Z3]; CVC3; veriT. \\
UFIDL & 	3 & 	49 & 	[Z3]; \win{CVC4}; CVC3. \\
UFLIA & 	4 & 	5766 & 	\win{CVC4}; [Z3]; veriT; CVC3. \\
UFLRA & 	4 & 	25 & 	[Z3]; \win{veriT}; CVC3; CVC4. \\
UFNIA & 	3 & 	1587 & 	[Z3]; \win{CVC4}; CVC3. \\
\hline
\end{tabular}}
\end{table}

>>>>

The results of the competition per se were not surprising. The solvers that performed best in previous years continued to do so. Boolector wins those divisions that focus on bit-vector problems; CVC4 has the most breadth of application and wins most of the other divisions, with Yices2 also having a strong showing. Solvers that were optimized with particular decision procedures for particular logics can capture individual wins: AProVE for QF\_NIA; veriT for UFLRA. If Z3 had competed, with the current version entered by the organizers, it would have won many divisions, but not all; CVC4 is 
definitely competitive with it.

@@

\begin{table}
\caption{Main track results}
\vspace{-\smallskipamount}
\label{Table:maintrack}
\centering
\newcommand{\win}[1]{\textbf{#1}}
\resizebox{!}{299pt}{\begin{tabular}{|l|r|r|l|}
\hline
Logic &	Solvers &	Benchmarks & Order (winner in bold) \\
\hline
ALIA & 		4 & 	29  & 	[Z3]; \win{CVC4}; veriT; CVC3. \\
AUFLIA & 	4 & 	4   & 	\win{CVC4}; [Z3]; CVC3; veriT. \\
AUFLIRA & 	4 & 	10791 & 	[Z3]; \win{CVC4}; CVC3; veriT. \\
AUFNIRA & 	4 & 	564 & 	[CVC4-with-bugfix]; [Z3]; \win{CVC3}; CVC4. \\
LIA & 		4 & 	46 & 	[Z3]; \win{CVC4}; CVC3; veriT. \\
LRA & 		4 & 	171 & 	\win{CVC4}; [Z3]; CVC3; veriT. \\
NIA & 		3 & 	9 &  	[Z3]; \win{CVC4}; CVC3. \\
NRA & 		3 & 	3747 & 	[Z3]; \win{CVC4}; CVC3. \\
QF\_ABV & 	9 & 	6457 & \win{Boolector-j}; Boolector-d; [MathSAT]; \\
& & & SONOLAR; CVC4; [Z3]; Yices2; \\
& & & Kleaver-STP; Kleaver-portfolio. \\
QF\_ALIA & 	5 & 	97 & \win{Yices2}; SMTInterpol; [Z3]; [MathSAT]; CVC4. \\
QF\_AUFBV & 	4 & 	37 & \win{CVC4}; Yices2; [Z3]; [MathSAT]. \\
QF\_AUFLIA & 	6 & 	610 & \win{Yices2}; [MathSAT]; [Z3] \\
& & & SMTInterpol; CVC4; veriT. \\
QF\_AX & 	5 & 	335 & \win{Yices2}; [MathSAT]; [Z3]; CVC4; \\
& & & SMTInterpol. \\
QF\_BV & 	11 & 	2488 & \win{Boolector}; STP-CryptoMiniSat4; \\
& & & [CVC4-with-bugfix]; [MathSAT]; [Z3]; \\
& & & CVC4; 4Simp; SONOLAR; Yices2; \\
& & & abziz\_min\_features; abziz\_all\_features. \\
QF\_IDL & 	4 & 	1315 & 	[Z3]; \win{Yices2}; CVC4; veriT. \\
QF\_LIA & 	7 & 	4381 & 	[CVC4-with-bugfix]; [MathSAT]; \\
& & & \win{SMTInterpol}; Yices2; [Z3]; veriT; CVC4. \\
QF\_LRA & 	6 & 	1343  & \win{CVC4}; Yices2; [MathSAT]; SMTInterpol; \\
& & & veriT; [Z3]. \\
QF\_NIA & 	4 & 	8327 & 	[Z3]; \win{AProVE}; CVC3; CVC4. \\
QF\_NRA & 	4 & 	10121 & [Z3]; \win{CVC3}; CVC4; raSAT. \\
QF\_RDL & 	4 & 	132 & 	\win{Yices2}; [Z3]; veriT; CVC4. \\
QF\_UF & 	7 & 	4124 & 	\win{Yices2}; veriT; CVC4; OpenSMT2; [Z3]; \\
& & & [MathSAT]; SMTInterpol. \\
QF\_UFBV & 	4 & 	31 & 	\win{Yices2}; [Z3]; [MathSAT]; CVC4. \\
QF\_UFIDL & 	4 & 	311 & 	[Z3]; \win{Yices2}; CVC4; veriT. \\
QF\_UFLIA & 	6 & 	484 & 	[Z3]; \win{Yices2}; CVC4; SMTInterpol; \\
& & & [MathSAT]; veriT. \\
QF\_UFLRA & 	6 & 	1176 & 	[Z3]; \win{Yices2}; [MathSAT]; CVC4; \\
& & & SMTInterpol; veriT. \\
QF\_UFNIA & 	3 & 	7 & 	\win{CVC4}; [Z3]; CVC3. \\
QF\_UFNRA & 	3 & 	32 & 	[Z3]; \win{CVC3}; CVC4. \\
UF & 		4 & 	2830 & 	\win{CVC4}; [Z3]; CVC3; veriT. \\
UFIDL & 	3 & 	49 & 	[Z3]; \win{CVC4}; CVC3. \\
UFLIA & 	4 & 	5766 & 	\win{CVC4}; [Z3]; veriT; CVC3. \\
UFLRA & 	4 & 	25 & 	[Z3]; \win{veriT}; CVC3; CVC4. \\
UFNIA & 	3 & 	1587 & 	[Z3]; \win{CVC4}; CVC3. \\
\hline
\end{tabular}}
\end{table}

\paragraph{Fastest solvers.} The competition does not weight time to solve a problem significantly. However, this is an important characteristic to any user. Hence we analyzed the competition results for each division to answer this question: 
for each solver, what fraction of the benchmarks for the division does that solver solve the fastest? 
The data answering this question is shown in Table \ref{tab:speed}. For each division and each benchmark, we determined the solver that solved that benchmark the fastest, and for each solver counted the number of benchmarks for which it was the fastest. The table lists the three solvers with the most such benchmarks, with the fraction of benchmarks that it won.
The winning solvers are a mix of Z3, CVC, Yices2, and a few for veriT. In some division, the winning solver does dominate teh competition. Yices2, for example tends to win in those divisions in which it participates, with Z3 winning most of the others. Note that this analysis does not take into account the margin by which the times are better. The times used here are cpu-times; usinjg wall-clock times changes the fractions and sometimes the order, but not the overall observation.

\begin{table}
\begin{tabular}{|l|lll|}
\hline
Logic & Fraction solved fastest & \\
\hline
ALIA & 51\% [Z3] & 48\% veriT  & \\
AUFLIA & 66\% [Z3] & 33\% CVC3  & \\
AUFLIRA & 60\% [Z3] & 37\% veriT & 1\% CVC3 \\
AUFNIRA & 72\% [Z3] & 23\% CVC3 & 2\% CVC4 \\
LIA & 43\% CVC4 & 30\% [Z3] & 17\% CVC3 \\
LRA & 42\% CVC4 & 38\% veriT & 18\% [Z3] \\
NIA & 66\% [Z3] & 33\% CVC4  & \\
NRA & 96\% CVC4 & 3\% [Z3]  & \\
QF\_ABV & 78\% Yices2 & 9\% SONOLAR & 7\% Kleaver-STP \\
QF\_ALIA & 95\% Yices2 & 4\% [Z3] & \\
QF\_AUFBV & 81\% Yices2 & 10\% [MathSAT] & 5\% [Z3] \\
QF\_AUFLIA & 60\% Yices2 & 38\% [Z3] & 1\% CVC4 \\
QF\_AX & 92\% Yices2 & 6\% [Z3] & 0\% CVC4 \\
QF\_BV & 33\% Yices2 & 25\% [Z3] & 13\% 4Simp \\
QF\_IDL & 87\% Yices2 & 11\% [Z3] & 0\% veriT \\
QF\_LIA & 76\% Yices2 & 11\% [Z3] & 7\% CVC4 \\
QF\_LRA & 86\% Yices2 & 3\% veriT & 3\% CVC4 \\
QF\_NIA & 90\% [Z3] & 7\% AProVE & 1\% CVC3 \\
QF\_NRA & 88\% [Z3] & 6\% CVC4 & 4\% CVC3 \\
QF\_RDL & 71\% Yices2 & 25\% [Z3] & 3\% veriT \\
QF\_UF & 96\% Yices2 & 1\% veriT & 1\% [Z3] \\
QF\_UFBV & 96\% Yices2 & 3\% [MathSAT]  & \\
QF\_UFIDL & 59\% [Z3] & 36\% Yices2 & 3\% veriT \\
QF\_UFLIA & 97\% Yices2 & 2\% [Z3] & 0\% CVC4 \\
QF\_UFLRA & 91\% Yices2 & 4\% [Z3] & 2\% veriT \\
QF\_UFNIA & 100\% CVC4 & \\
QF\_UFNRA & 59\% CVC3 & 37\% [Z3] & 3\% CVC4 \\
UF & 37\% veriT & 32\% CVC4 & 24\% [Z3] \\
UFIDL & 57\% CVC4 & 42\% [Z3]  & \\
UFLIA & 56\% [Z3] & 35\% veriT & 7\% CVC4 \\
UFLRA & 95\% veriT & 4\% [Z3]  & \\
UFNIA & 80\% [Z3] & 10\% CVC3 & 9\% CVC4 \\

\hline
\end{tabular}
\caption{Fraction of benchmarks solved fastest (cpu-time) by each solver (top 3)}
\label{tab:speed}
\end{table}

\paragraph{Solver contribution.} A second comparison that can be made is to determine what a solver adds in addition to 
what other solvers contribute. We measure this `solver contribution' as follows. For each benchmark in a division, we award each solver that solves the benchmarks $1/n$ points, where $n$ is the number of solvers that solved the benchmark (within the timeout). These points are then summed over all the benchmarks in the division and scaled by the number of benchmarks. The sum is a measure of the
unique contribution provided by a solver; a value of 1.0 would indicate that that solver was the only solver to solve all of the benchmarks in the division. The results are shown in Table \ref{tab:cont}. The results show again that no one solver dominates in all categories, but that there are several contributing solvers among which users can choose.

\begin{table}
\begin{tabular}{|l|lll|}
\hline
Logic & Unique contribution & & \\
\hline
ALIA & 0.4253 [Z3] & 0.4080 CVC4 & 0.1667 veriT \\
AUFLIA & 0.3889 CVC4 & 0.2778 [Z3] & 0.2222 CVC3 \\
AUFLIRA & 0.2615 [Z3] & 0.2521 CVC4 & 0.2503 CVC3 \\
AUFNIRA & 0.2509 [Z3] & 0.2506 CVC4 & 0.2497 CVC3 \\
LIA & 0.3949 [Z3] & 0.3514 CVC4 & 0.2319 CVC3 \\
LRA & 0.3070 CVC4 & 0.3070 [Z3] & 0.2895 CVC3 \\
NIA & 0.8333 [Z3] & 0.1667 CVC4 & 0.0000 Yices2 \\
NRA & 0.5071 [Z3] & 0.4929 CVC4 & 0.0000 Yices2 \\
QF\_ABV & 0.1147 Yices2 & 0.1146 Boolector-just... & 0.1144 Boolector-dual... \\
QF\_ALIA & 0.2235 Yices2 & 0.2235 SMTInterpol & 0.2235 [Z3] \\
QF\_AUFBV & 0.2635 [MathSAT] & 0.2635 CVC4 & 0.2500 Yices2 \\
QF\_AUFLIA & 0.1996 Yices2 & 0.1996 [MathSAT] & 0.1996 CVC4 \\
QF\_AX & 0.2000 Yices2 & 0.2000 [MathSAT] & 0.2000 CVC4 \\
QF\_BV & 0.1051 Boolector & 0.0976 STP-Crypto... & 0.0934 abziz\_min\_features \\
QF\_IDL & 0.2893 [Z3] & 0.2731 Yices2 & 0.2395 CVC4 \\
QF\_LIA & 0.1620 [CVC4-with-bugfix] & 0.1610 [Z3] & 0.1610 SMTInterpol \\
QF\_LRA & 0.1772 CVC4 & 0.1689 Yices2 & 0.1678 SMTInterpol \\
QF\_NIA & 0.4990 [Z3] & 0.4912 AProVE & 0.0084 CVC3 \\
QF\_NRA & 0.7714 [Z3] & 0.1347 CVC3 & 0.0898 CVC4 \\
QF\_RDL & 0.2734 Yices2 & 0.2431 veriT & 0.2431 [Z3] \\
QF\_UF & 0.1451 Yices2 & 0.1449 veriT & 0.1446 CVC4 \\
QF\_UFBV & 0.4140 Yices2 & 0.3172 [Z3] & 0.1720 [MathSAT] \\
QF\_UFIDL & 0.2708 Yices2 & 0.2708 [Z3] & 0.2309 CVC4 \\
QF\_UFLIA & 0.1698 Yices2 & 0.1698 [MathSAT] & 0.1698 CVC4 \\
QF\_UFLRA & 0.1715 [Z3] & 0.1681 Yices2 & 0.1681 [MathSAT] \\
QF\_UFNIA & 0.5000 CVC4 & 0.5000 [Z3] & 0.0000 Yices2 \\
QF\_UFNRA & 0.5802 [Z3] & 0.2099 CVC4 & 0.2099 CVC3 \\
UF & 0.4921 CVC4 & 0.1988 [Z3] & 0.1759 CVC3 \\
UFIDL & 0.5204 [Z3] & 0.4796 CVC4 & 0.0000 Yices2 \\
UFLIA & 0.4069 CVC4 & 0.3502 [Z3] & 0.2425 veriT \\
UFLRA & 0.2857 [Z3] & 0.2381 CVC4 & 0.2381 veriT \\
UFNIA & 0.4156 [Z3] & 0.3860 CVC4 & 0.1984 CVC3 \\
\hline
\end{tabular}
\caption{Contribution from each solver (top 3)}
\label{tab:cont}
\end{table}
